---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: Nora Jones 
    url: https://example.com/norajones
    affiliation: Spacely Sprockets
    affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web. 

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.



```{r}

packages = c('tmap', 'sf', 'readr',
             'readxl', 'knitr',
             'data.table', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}

```

```{r}
sgpools <- read_csv("data/SGPools_svy21.csv")
```



```{r}
sgpools_sf <- st_as_sf(
  sgpools, 
  coords = c("XCOORD",
             "YCOORD"),
             crs= 3414)
```

tm_shape defines the extent of the outter layer
border.lwd:line size
```{r}
tmap_mode("view")
tm_shape(sgpools_sf)+
tm_bubbles(col = "red",
           size = 1,
           border.col = "black",
           border.lwd = 1)
```


```{r}
tmap_mode("view")
tm_shape(sgpools_sf)+
tm_bubbles(col = "red",
           size = 1,
           border.col = "black",
           border.lwd = 1)
```




```{r}
tm_shape(sgpools_sf)+
tm_bubbles(col = "OUTLET TYPE",
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 0.5)
```


```{r}
tm_shape(sgpools_sf) +
  tm_bubbles(col = "OUTLET TYPE", 
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 1) +
  tm_facets(by= "OUTLET TYPE",
            nrow = 1,
            sync = TRUE)
```
```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```


```{r}
popagsex <- read_csv("data/respopagsex2000to2018.csv")
```
spread ==> pivot_wide;
`0_to_4`+`5_to_9`+`10_to_14`+`15_to_19`+`20_to_24` ==> rowsums(.[5:9])

```{r}
popagsex2018_male <- popagsex %>%
  filter(Sex == "Males") %>%
  filter(Time == 2018) %>%
  pivot_wider(names_from = AG, values_from = Pop) %>%
  mutate(YOUNG = `0_to_4`+`5_to_9`+`10_to_14`+
`15_to_19`+`20_to_24`) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[9:13])+
rowSums(.[15:17]))%>%
mutate(`AGED`=rowSums(.[18:22])) %>%
mutate(`TOTAL`=rowSums(.[5:22])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  select(`PA`, `SZ`, `YOUNG`, 
         `ECONOMY ACTIVE`, `AGED`, 
         `TOTAL`, `DEPENDENCY`) %>%
  filter(`ECONOMY ACTIVE` > 0)
```


```{r}
mpsz_agemale2018 <- left_join(mpsz,
                              popagsex2018_male,
                              by = c("SUBZONE_N" = "SZ"))
```


```{r}
tmap_mode("plot")
qtm(mpsz_agemale2018, 
    fill = "DEPENDENCY")
```
```{r}
tm_shape(mpsz_agemale2018) +
  tm_polygons()
```

```{r}
tm_polygons("DEPENDENCY")
```

```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill("DEPENDENCY")
```


```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,
             alpha = 1)
```


style = "quantile" clustering by quantile, or clustering, etc. n=8 ==> 8 bins

```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill("DEPENDENCY",
          n = 8,
          style = "quantile") +
  tm_borders(alpha = 0.5)
```
style = "equal"

```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill("DEPENDENCY",
          n = 8,
          style = "equal") +
  tm_borders(alpha = 0.5)
```


```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```
To reverse the colour shading, add a "-" prefix.
```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Blues") +
  tm_borders(alpha = 0.5)
```

Map Layouts

```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Quantile classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45,
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```



```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```



```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS",
             position = c("left", "bottom"))+
  tmap_style("white")
```

```{r}
tm_shape(mpsz_agemale2018)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right",
                                "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```
free.coords=TRUE to have different shape form different regions individually 
drop.shapes=TRUE neighbouring regions to have blank shape

```{r}
tm_shape(mpsz_agemale2018) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N",
            free.coords=TRUE,
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center",
                               "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```
```{r}
youngmap <- tm_shape(mpsz_agemale2018)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")
agedmap <- tm_shape(mpsz_agemale2018)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")
tmap_arrange(youngmap,
             agedmap,
             asp=1,
             ncol=2)
```

```{r}
tm_shape(mpsz_agemale2018[mpsz_agemale2018$REGION_N=="CENTRAL REGION", ]) +
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```
