---
title: "In-class_Ex01 Building Elegant Graphics with ggplot2"
description: |

author:
  - name: Sun Shengmei
    url: https://www.linkedin.com/in/shengmei-sun-9b262656/?originalSubdomain=sg
    affiliation: SMU MITB
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "`r Sys.Date()`"
output: distill::distill_article
---

# echo = True: to display the code on the webpage
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.retina = 3)
```

# To install packages: 'tidyverse','ggrepel'
```{r}
packages = c('tidyverse','ggrepel')
for(p in packages){
   if(!require(p, character.only = T)){
     install.packages(p)
   }
   library(p, character.only = T)
}
```

# To import data using read_csv of readr package

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

```{r}
ggplot(data=exam_data,
       aes(x=RACE)) +
  geom_bar()
```
# "scale_y_continuous" is to turn off the y-axis
```{r}
ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_dotplot(binwidth= 2.5,
               dotsize = 0.5) +
  scale_y_continuous(NULL,
                     breaks = NULL)
```

```{r}
ggplot(data=exam_data,
       aes(x = MATHS)) +
  geom_histogram()

```
# "Fill" inside the geom_histogram, give the same color to all the bars. "color" is for the border
```{r}
ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_histogram(bins=20,
                 color="black",
                 fill="light blue") +
  ggtitle("Distribution of Maths scores")
```
# Staked histogram. "Color" here is for the outline of the bar. "fill" inside the aes gives 2 colors by gender.
```{r}
ggplot(data=exam_data,
       aes(x=MATHS, 
           fill= GENDER)) +
  geom_histogram(bins=20,
                 color="grey30")
```
```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           color=GENDER)) +
  geom_density()
```
# "alpha" is to define the tickness of the line. Boxplot shows median VS lower 25% VS upper 25%.
```{r}
ggplot(data=exam_data,
       aes(y=MATHS, 
           x= GENDER)) +
  geom_violin(fill="lightblue")+
  geom_boxplot(alpha=0.5)
```
```{r}
ggplot(data=exam_data,
       aes(x = MATHS, y=ENGLISH)) +
  geom_point()
```
# To add in summary statistics/mean method 1
```{r}
ggplot(data=exam_data,
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot(alpha=0.5) +
  stat_summary(geom = "point",
               fun.y = "mean",
               color ="red",
               size=4)
```
# To add in summary statistics/mean method 2
```{r}
ggplot(data=exam_data,
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot(alpha=0.5) +
  geom_point(stat = "summary",
             fun.y= "mean",
             color ="red",
             size=4)

```
```{r}
ggplot(data=exam_data,
       aes(x = MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size = 0.5)
```
```{r}
ggplot(data=exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins=20)+
  facet_wrap(~CLASS)
```
```{r}
ggplot(data=exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins=20)+
  facet_grid(~CLASS)
```
```{r}
ggplot(data=exam_data,
       aes(y = MATHS, x=CLASS)) +
  geom_boxplot()+
  facet_grid(~GENDER)
```
```{r}
ggplot(data=exam_data,
       aes(y = MATHS, x=CLASS)) +
  geom_boxplot()+
  facet_grid(GENDER~.)
```
```{r}
ggplot(data=exam_data,
       aes(y = MATHS, x=GENDER)) +
  geom_boxplot()+
  facet_grid(GENDER~CLASS)
```
```{r}
ggplot(data=exam_data,
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()
```
```{r}
ggplot(data=exam_data,
       aes(x=MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
```
```{r}
ggplot(data=exam_data,
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()+
  theme_gray()
```
```{r}
ggplot(data=exam_data,
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()+
  theme_classic()
```
```{r}
ggplot(data=exam_data,
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()+
  theme_minimal()
```
```{r}
ggplot(data=exam_data,
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()+
  theme_minimal()+
  theme(panel.background = element_rect(fill="lightblue",
                                       color = "lightblue",
                                       size=0.5,
                                       linetype="solid"),
        panel.grid.major = element_line(size = 0.5, 
                                        linetype = "solid",
                                        color = "white"),
        panel.grid.minor = element_line(size = 0.25, 
                                        linetype = "solid",
                                        color = "white"))
```
# 不明白这个function怎么work. “geom_text" is used for labelling plots
# You can modify text alignment with the vjust and hjust aesthetics. These can either be a number between 0 (right/bottom) and 1 (top/left) or a character ("left", "middle", "right", "bottom", "center", "top"). There are two special alignments: "inward" and "outward". Inward always aligns text towards the center, and outward aligns it away from the center.
```{r}
ggplot(data=exam_data,
       aes(x=reorder(RACE, RACE, 
                     function(x) - length(x)))) +
  geom_bar()+
  ylim(0,220)+
  geom_text(stat="count",
            aes(label=paste0(..count..,",",
            round(..count../sum(..count..)*100,
                  1),"%")),
            vjust=-1)+
  xlab("Race")+
  ylab("No.of\nPupils")+
  theme(axis.title.y=element_text(angle=0))
         
```
# %>%：to avoid nesting loop? 不明白
# mutate: add new variables and preserve existing ones
```{r}
exam_data %>%
  mutate(RACE=fct_infreq(RACE)) %>%
  ggplot(aes(x=RACE)) +
  geom_bar()+
  ylim(0,220)+
  geom_text(stat="count",
            aes(label=paste0(..count..,",",
            round(..count../sum(..count..)*100,
                  1),"%")),
            vjust=-1)+
  xlab("Race")+
  ylab("No.of\nPupils")+
  theme(axis.title.y=element_text(angle=0))
```
# geom_vline: to add reference line
# na.rm=T: na value will be stripped before computation
```{r}
ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_histogram(bins=20,
                 color="black",
                 fill="light blue") +
  geom_vline(aes(xintercept=mean(MATHS,na.rm=T)),
             color="red",
             linetype="dashed",
             size=1)+
  geom_vline(aes(xintercept=median(MATHS,na.rm=T)),
             color="grey30",
             linetype="dashed",
             size=1)
```
# guides(fill=FALSE): no legend. if guides(): there is legend
```{r}
d <- exam_data
d_bg <- d[,-3]
ggplot(d, aes(x=ENGLISH,
              fill = GENDER))+
  geom_histogram(data = d_bg,
                 fill="grey",
                 alpha =.5)+
  geom_histogram(colour="black")+
  facet_wrap(~GENDER) +
  guides(fill = FALSE) +
  theme_bw()
```
```{r}
ggplot(data=exam_data,
       aes(x=MATHS, y=ENGLISH))+
  geom_point()+
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))+
  geom_hline(yintercept=50,
             linetype="dashed",
             color="grey60",
             size=1)+
  geom_vline(xintercept=50,
             linetype="dashed",
             color="grey60",
             size=1)
```






